---
title: "Week 2 Exercises"
author: "PJ Grant"
date: "October 27, 2024"
output: pdf_document
---

Please complete all exercises below. You may use stringr, lubridate, or the forcats library. 

Place this at the top of your script:
library(stringr)
library(lubridate)
library(forcats)

```{r}
library(stringr)
library(lubridate)
library(forcats)
```



# Exercise 1
Read the sales_pipe.txt file into an R data frame as sales. 

```{r}
sales <- read_delim("~/MM/DSE5002/Week_2/Data/sales_pipe.txt", 
                    delim = "|", 
                    escape_double = FALSE, 
                    trim_ws = TRUE, 
                    locale = locale(encoding = "latin1"))

# Splitting the values in the profit column into three distinct columns
temp_char <- str_split_fixed(string=sales$Profit,pattern='\\|', n=3)

#Assign the correct column headers
colnames(temp_char) <- c("Quantity1", "Discount1", "Profit1")

#Combining the two data frames
updated_sales <- cbind(sales, temp_char)   

# Super setting the correct columns
updated_sales <- updated_sales[,2:23]

#eliminating the incorrect profit column
updated_sales <- updated_sales[,-19]

#Renaming the columns
new_column_names <- c("row_id", "order_id", "order_date", "ship_date", "ship_mode", "customer_id", "customer_name", "segment", "country", "city", "state", "postal_code", "region", "product_id", "category", "sub_category", "product_name", "sales", "quantity", "discount", "profit")

colnames(updated_sales) <- new_column_names

#updating from strings / characters to numbers
updated_sales$quantity = as.numeric(updated_sales$quantity)
updated_sales$discount = as.numeric(updated_sales$discount)
updated_sales$profit = as.numeric(updated_sales$profit)

#convert back to sales
sales = updated_sales
View(sales)

```

# Exercise 2
You can extract a vector of columns names from a data frame using the colnames() function. Notice the first column has some odd characters. Change the column name for the FIRST column in the sales date frame to Row.ID. 

**Note: You will need to assign the first element of colnames to a single character.**

```{r}
# Included in the column name update above
```

# Exercise 3
Convert both Ship.Date and Order.Date to date vectors within the sales data frame. What is the number of days between the most recent order and the oldest order?
How many years is that?
How many weeks?

**Note: Use lubridate**

```{r}
#Set as date
sales$ship_date <- as.Date(sales$ship_date, format = '%B %d %Y')
sales$order_date <- as.Date(sales$order_date, format = '%m/%d/%Y')

#Find max and min order dates
newest_order = max(sales$order_date) #The most recent order date is 2017-12-30
oldest_order = min(sales$order_date) #The older order was placed on 2014-01-03

#Create interval between oldest and most recent order
date_intervals = interval(newest_order, oldest_order)

time_in_seconds = int_length(date_intervals) * -1
time_in_seconds


#Convert interval from seconds to days // 1457 days occured between the oldest and most recent order
number_of_days = time_in_seconds / 60 / 60 / 24
number_of_days

# ~ 4 years
number_of_days / 365

# ~208 weeks
number_of_days / 7

```


# Exercise 4
What is the average number of days it takes to ship an order?

```{r}
#Calculate the time to ship for each order
sales$time_to_ship <- sales$ship_date - sales$order_date

# The average time to ship for each order is 3.9 days
mean(sales$time_to_ship)

```


# Exercise 5
How many customers have the first name Bill?
You will need to split the customer name into first and last name segments and then use a regular expression to match the first name bill. Use the length() function to determine the number of customers with the first name Bill in the sales data. 

```{r}
#Split names and create new columns with first and last name
sales$name <- str_split_fixed(sales$customer_name," ", n=2)
sales$first_name <- sales$name[,1]
sales$last_name<- sales$name[,2]

# There are 37 people names bill
length(str_subset(sales$first_name, "Bill"))

```

# Exercise 6
How many mentions of the word 'table' are there in the Product.Name column? 
**Note you can do this in one line of code**

```{r}
#240 mentions of the word table in the product_name column
sum(str_count(sales$product_name,"table"))
```

# Exercise 7
Create a table of counts for each state in the sales data. The counts table should be ordered alphabetically from A to Z. 

```{r}
table(sales$state)
```

# Exercise 8
Create an alphabetically ordered barplot for each sales Category in the State of Texas. 

```{r}
texas_count <- sales[sales$state == "Texas",]
barplot(table(texas_count$category))
```

# Exercise 9
Find the average profit by region.
**Note: You will need to use the aggregate() function to do this. To understand how the function works type ?aggregate in the console.**

```{r}
#The average profit by region is as follows: Central(20.5), East(29.9), South(11.3), West(32.8) 
aggregate(sales$profit, list(sales$region), mean)
```

# Exercise 10
Find the average profit by order year. 
**Note: You will need to use the aggregate() function to do this. To understand how the function works type ?aggregate in the console.**

```{r}
#Determine which year each order was placed
sales$order_year <- year(sales$order_date)

#Determine the average profit by year
aggregate(sales$profit, list(sales$order_year), mean)

#The average profit for 2014 was 32.2, 2015 was 21.6, 2016 was 30.1, 2017 was 21.3 
```

